---
title: "Data"
output: 
  learnr::tutorial:
   progressive: true
   allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}

library(learnr)
knitr::opts_chunk$set(echo = TRUE)
learnr::tutorial_options(exercise.completion = FALSE)

```

## Disclaimer

The content of this tutorial draws heavily on the following sources:

- [R4DS by Garrett Grolemund & Hadley Wickham](https://r4ds.had.co.nz/)

- [Advanced R by Hadley Wickham](https://adv-r.hadley.nz/index.html#license)

Before you proceed with this tutorial, make sure you have completed the
following Rstudio primers:

1. [The basics](https://rstudio.cloud/learn/primers/1)

1. [Work with data](https://rstudio.cloud/learn/primers/2)


## Introduction

This is the third tutorial in a series that I have created to help 
you learn more about R. Since this tutorial focuses on data, let's
start with a broad definition of the `data.frame` class:

### What is a `data.frame`?

A `data.frame` is a special type of `list` vector. Specifically, a `data.frame`
is a `list` that has the following conditions:

- All elements of the `data.frame` are vectors comprising data of one type.

- All elements of the `data.frame` have equal length.

- All elements of the `data.frame` have a unique name.

### Creating a `data.frame`

Since a `data.frame` is a special kind of list, it is possible to coerce
certain lists into a `data.frame`:

```{r}

lst <- list(
  nmrs = 1:5,
  chrs = letters[1:5]
)

df <- as.data.frame(lst)

str(df)

```

You can also use the `data.frame()` function.

```{r}

df <- data.frame(
  nmrs = 1:5,
  chrs = letters[1:5]
)

str(df)

```

You can access attributes of `data.frame` objects just like vectors:

```{r}

# Print the column names
names(df)
colnames(df)

# Print the rownames
rownames(df)

# Print the number of rows and number of columns
dim(df)

# Print just the number of rows
nrow(df)

# Print just the number of columns
ncol(df)

```

### What if inputs are wrong?

R will go to great lengths to help you convert a `list` into a `data.frame`.

- If the component vectors are not of equal length, R will try to correct
  the inputs using vector recycling.
  
  ```{r error = TRUE}
  
  # Recycling works here:
  lst <- list(nmrs = 1, chrs = letters[1:5])
  df <- as.data.frame(lst)
  str(df)
  
  # And here...
  lst <- list(nmrs = 1:3, chrs = letters[1:6])
  df <- as.data.frame(lst)
  str(df)
  
  # Will recycling work here?
  lst <- list(nmrs = 1:3, chrs = letters[1:5])
  df <- as.data.frame(lst)
  
  
  ```
  
- If the components have data of a different type, R will try to coerce 
  them into vectors containing data of a single type.

  ```{r}
  
  lst <- list(nmrs = 1:5, chrs = list(lst_chr = "A", lst_nmr = 1))
  df <- as.data.frame(lst)
  str(df)  
  
  ```

- If the vectors are not named, R will name them (poorly).

```{r}

lst <- list(1:5, letters[1:5])
df <- as.data.frame(lst)
str(df)

```

### Your turn

Print a data frame with the following names and values

- Name: `id`, values: `1:100`

- Name: `score`, values: `100`

- Name: `type`, values: the letter `'A'` repeated 50 times, followed 
  by the letter `'B'` repeated 50 times.

```{r make_df, exercise = TRUE}



```

```{r make_df-solution}

data.frame(
  id = 1:100,
  score = 100,
  type = rep(c("A","B"), each = 50)
)

```

## Tidy tables (`tibble`)

A `tibble` is an extension of the standard R `data.frame` class. In other 
words, `tibble`s _are_ data frames, but they tweak some older behaviours 
to make life a little (sometimes a lot) easier. R is an old language, and 
some things that were useful 10 or 20 years ago now get in your way. It's
difficult to change base R without breaking existing code, so most innovation
occurs in packages. Here we will describe the __tibble__ package, which 
provides opinionated data frames that make data analysis a little easier. 
From here on out, I'll use the term `tibble` and data frame interchangeably; 
when I want to draw particular attention to the standard R data frame, I'll 
use the term `data.frame`. 

If this chapter leaves you wanting to learn more about tibbles, you might 
enjoy reading more at [tidyverse.org](https://tibble.tidyverse.org/index.html)

### Creating a `tibble`

A `data.frame` can be converted into a `tibble` using `as_tibble()`:

```{r}

# load the tibble package to access tibble functions
library(tibble)

df <- data.frame(
  nmrs = 1:5,
  chrs = letters[1:5]
)

as_tibble(df)

```

A `tibble` can also be created directly using the `tibble()` function.

```{r}

tibble_df <- tibble(
  nmrs = 1:5,
  chrs = letters[1:5]
)

tibble_df

```

`tibble()` will automatically recycle inputs of length 1 and allow you to 
refer to variables that you just created, as shown below.

```{r}

tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)

```

`tibble()` does much less than `data.frame()`. it never changes the type of 
the inputs (e.g. it never converts strings to factors!), it never changes 
the names of variables, and it never creates row names.

### Non-syntactic names

It's possible for a tibble to have column names that are not valid R variable
names, aka __non-syntactic__ names. For example, they might not start with a
letter, or they might contain unusual characters like a space. To refer to 
these variables, you need to surround them with backticks, `` ` ``:

```{r}
tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb
```

You'll also need the backticks when working with these variables in other packages, like `ggplot2`, `dplyr`, and `tidyr`.

## `tibble` vs. `data.frame`

There are two main differences in the usage of a `tibble` vs. a classic `data.frame`: printing and subsetting.

### Printing

Tibbles have a refined print method that shows only the first 10 rows, and all the columns that fit on screen. This makes it much easier to work with large data. In addition to its name, each column reports its type, a nice feature borrowed from `str()`:

```{r}

tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)

```

Tibbles are designed so that you don't accidentally overwhelm your console 
when you print large data frames. But sometimes you need more output than 
the default display. There are a few options that can help.

First, you can explicitly `print()` the data frame and control the number 
of rows (`n`) and the `width` of the display. `width = Inf` will display 
all columns:

```{r, eval = FALSE}
nycflights13::flights %>% 
  print(n = 10, width = Inf)
```

You can also control the default print behaviour by setting options:

* `options(tibble.print_max = n, tibble.print_min = m)`: if more than `n`
  rows, print only `m` rows. Use `options(tibble.print_min = Inf)` to always
  show all rows.

* Use `options(tibble.width = Inf)` to always print all columns, regardless
  of the width of the screen.

You can see a complete list of options by looking at the package help with `package?tibble`.

A final option is to use RStudio's built-in data viewer to get a scrollable view of the complete dataset. This is also often useful at the end of a long chain of manipulations.

```{r, eval = FALSE}
nycflights13::flights %>% 
  View()
```

